Regional Inequalities and economic growth; cross section and times series analysis

### 2023-12-06 MSB104_Submission_Group1.qmd
M

Why some places grow and prosper is a fundamental question in the field of social science. This question motivated Adam Smith's \[1863\] work and has been a major influence of the *new economic geography* [@feldman2014]*. Furthermore,* regional growth and inequality have been important topics of research in Europe for many years. Despite the European commission\'s efforts to decrease income gaps, significant disparities in income and economic development persist across regions in Europe [@lessmann2017].

@@

Regional growth in Europe reveals a complex pattern, influenced significantly by factors such as geography, human capital, and economic policies. Wealthier nations like Denmark tend to experience quicker regional convergence, a phenomenon supported by \[@gennaioli2014\] findings.

@@

The effect that surprised us most was that the economic crisis led to increase inequality\[@nguyen2022\]. indicate that any type of financial crisis results in higher income inequality. Furthermore,  @lessmann2017 address the importance of studying these aspects. Moreover, its potential consequences, such as political tensions that can undermine social and political stability [@lessmann2017].

@@

All dataset used in this paper is from Eurostat and based on the Nomenclature of Territorial Units for Statistics (NUTS) classification. The NUTS system provides a hierarchical classification of the EU\'s economic territory, facilitating the collection and harmonization of European statistics, and ranges from the national level (NUTS 0) to smaller regional divisions (NUTS 3).

@@

Heteroskedasticity is according to @wooldridge2020 an important factor to consider in econometrics and statistical modelling, especially in the context of regression analysis. Heteroskedasticity refers to the condition in regression analysis where the variance of the error terms (residuals) is not constant across all levels of the independent variables \[@wooldridge2020\]. In simpler terms, the spread of the residuals varies at different points in the regression model. Moreover, heteroskedasticity can lead to inefficient estimators and unreliable hypothesis tests. Leading to the assumption of homoscedastic in MRL not being met. In other words the variance of the error terms is not constant across all levels if the independent variables when heteroskedasticity is present.

@@

The residuals versus fitted values plot is another diagnostic tool that helps in visually identifying issues in a regression model, aiding in improving model accuracy and reliability [@wooldridge2020]. Even after confirming heteroskedasticity. According to @wooldridge2020 It provides a deeper understanding of the heteroskedasticity's nature and extent, which is crucial for effectively addressing it in our regression analysis. Furthermore, the interpretation of different patterns in the plot, like random dispersion indicating a good model fit or systematic patterns indicating issues. Subsiquently, ensuring robust and reliable modeling visual diagnostics, complement statistical tests \[@wooldridge2020\].

>>>>

Why some places grow and prosper is a fundamental question in the field of social science. This question motivated Adam Smith's \[1863\] work and has been a major influence of the *new economic geography* [@feldman2014]*. Furthermore,* regional growth and inequality have been important topics of research in Europe for many years. Despite the European commission's efforts to decrease income gaps, significant disparities in income and economic development persist across regions in Europe [@lessmann2017].

@@

Regional growth in Europe reveals a complex pattern, influenced significantly by factors such as geography, human capital, and economic policies. Wealthier nations like Denmark tend to experience quicker regional convergence, a phenomenon supported by [@gennaioli2014] findings.

@@

The effect that surprised us most was that the economic crisis led to increase inequality[@nguyen2022]. indicate that any type of financial crisis results in higher income inequality. Furthermore,  @lessmann2017 address the importance of studying these aspects. Moreover, its potential consequences, such as political tensions that can undermine social and political stability [@lessmann2017].

@@

All dataset used in this paper is from Eurostat and based on the Nomenclature of Territorial Units for Statistics (NUTS) classification. The NUTS system provides a hierarchical classification of the EU's economic territory, facilitating the collection and harmonization of European statistics, and ranges from the national level (NUTS 0) to smaller regional divisions (NUTS 3).

@@

Heteroskedasticity is according to @wooldridge2020 an important factor to consider in econometrics and statistical modelling, especially in the context of regression analysis. Heteroskedasticity refers to the condition in regression analysis where the variance of the error terms (residuals) is not constant across all levels of the independent variables [@wooldridge2020]. In simpler terms, the spread of the residuals varies at different points in the regression model. Moreover, heteroskedasticity can lead to inefficient estimators and unreliable hypothesis tests. Leading to the assumption of homoscedastic in MRL not being met. In other words the variance of the error terms is not constant across all levels if the independent variables when heteroskedasticity is present.

@@

The residuals versus fitted values plot is another diagnostic tool that helps in visually identifying issues in a regression model, aiding in improving model accuracy and reliability [@wooldridge2020]. Even after confirming heteroskedasticity. According to @wooldridge2020 It provides a deeper understanding of the heteroskedasticity's nature and extent, which is crucial for effectively addressing it in our regression analysis. Furthermore, the interpretation of different patterns in the plot, like random dispersion indicating a good model fit or systematic patterns indicating issues. Subsiquently, ensuring robust and reliable modeling visual diagnostics, complement statistical tests [@wooldridge2020].
